---
output:
  github_document:
    html_preview: true
  html_document:
    keep_md: yes
always_allow_html: yes
editor_options: 
  chunk_output_type: console
---

<!-- README.Rmd generates README.md. -->

```{r, echo=FALSE}
knitr::opts_chunk$set(
  # collapse = TRUE,
  fig.align = "center",
  comment = "#>",
  fig.path = "man/figures/",
  message = FALSE,
  warning = FALSE
)

options(width = 120)
```


# `{tweetio}`


<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/tweetio)](https://cran.r-project.org/package=tweetio)
[![Lifecycle](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![GitHub last commit](https://img.shields.io/github/last-commit/knapply/tweetio.svg)](https://github.com/knapply/tweetio/commits/master)
[![Codecov test coverage](https://codecov.io/gh/knapply/tweetio/branch/master/graph/badge.svg)](https://codecov.io/gh/knapply/tweetio?branch=master)
[![AppVeyor build status](https://ci.appveyor.com/api/projects/status/github/knapply/tweetio?branch=master&svg=true)](https://ci.appveyor.com/project/knapply/tweetio)
[![Travis-CI Build Status](https://travis-ci.org/knapply/tweetio.svg?branch=master)](https://travis-ci.org/knapply/tweetio)
[![License: GPL v3](https://img.shields.io/badge/License-GPLv3-blue.svg)](https://www.gnu.org/licenses/gpl-3.0)
[![Depends](https://img.shields.io/badge/Depends-GNU_R>=3.3-blue.svg)](https://www.r-project.org/)
[![GitHub code size in bytes](https://img.shields.io/github/languages/code-size/knapply/tweetio.svg)](https://github.com/knapply/tweetio)
[![HitCount](http://hits.dwyl.io/knapply/tweetio.svg)](http://hits.dwyl.io/knapply/tweetio)
[!GH-Stars](https://img.shields.io/github/stars/knapply/tweetio?style=social)
<!-- badges: end -->

# Progress

### Supported Data Inputs

- [x] Twitter API streams: .json, .json.gz
- [x] API to Elasticsearch data dump (JSON Array): .json, .json.gz
- [x] API to Elasticsearch data dump (line-delimited JSON): .jsonl, .jsonl.gz
- [x] CSV (if written via `tweetio::write_tweet_csv()`)


### Supported Data Outputs

- [x] CSV
- [x] Excel
- [ ] .graphml (currently available in [`{tweetgraph}`](https://knapply.github.io/tweetgraph/))

### Extras
- [x] Spatial Tweets via `tweetio::as_tweet_sf()`


# Introduction

`{tweetio}` (mainly) focuses on one thing: __going from raw tweets to `{rtweet}`-style data frames and other useful structures) as quickly as possible__. Whether the data came from the Twitter API, a database dump, or some other source, `{tweetio}`'s job is to get them into R.

`{tweetio}` is __not__ a competitor to `{rtweet}`; it is not interested in collecting Twitter data. That said, it definitely attempts to emulate its data frame schema because...

1. It's incredibly easy to use.
2. It's more efficient to analyze than a key-value format following the raw data.
3. It'd be a waste not to maximize compatibiltiy with tools built specifically around `{rtweet}`'s data frames.


# Installation

You'll need a C++ compiler. If you're using Windows, you'll need [Rtools](https://cran.r-project.org/bin/windows/Rtools/).

You probably want to follow the page's advice and select the recommended version, which is currently Rtools35.exe. When you're installing Rtools, you need to make sure you check the box stating “Add rtools to system PATH” and reboot when you're finished.

Once you're done, reboot. Then you can install `{tweetio}`.


```{r, eval=FALSE}
# install.packages("remotes")
remotes::install_github("knapply/tweetio")
```




# Usage

```{r}
library(tweetio)
```

## Simple Example

First, we'll save a stream of tweets using `rtweet::stream_tweets()`.

```{r}
temp_file <- tempfile(fileext = ".json")
rtweet::stream_tweets(timeout = 15, parse = FALSE, file_name = temp_file)
```

We can then pass the file path to `tweetio::read_tweets()` to efficiently parse the data into an `{rtweet}`-style data frame.

```{r}
small_rtweet_stream <- read_tweets(temp_file, as_tibble = TRUE)

small_rtweet_stream
```

## Scaling Up

We're more interested in handling much larger data sets, but for demonstration we'll use a file of a reasonable size that was obtained using `rtweet::stream_tweets()`. 

```{r}
rtweet_stream_path <- "inst/example-data/api-stream.json.gz"

scales::number_bytes(file.size(rtweet_stream_path))
```

Before compression, the file was about 153 Mb.

Unfortunately, `rtweet::parse_stream()` may fail parsing streams because the data returned may not be valid JSON.

```{r, error=TRUE}
rtweet::parse_stream(rtweet_stream_path)
```

The only way around this is to read the entire file into memory and run some validation routine before parsing, which looks something like this...

```{r}
rtweet_validate_and_parse <- function(source_file_path, target_file_path) {
  raw_lines <- readLines(source_file_path)
  valid_lines <- raw_lines[vapply(raw_lines, jsonify::validate_json, logical(1L))]
  writeLines(valid_lines, target_file_path)
  
  rtweet::parse_stream(target_file_path)
}

rtweet_stream_path2 <- tempfile(fileext = ".json")

rtweet_validate_and_parse(source_file_path = rtweet_stream_path, 
                          target_file_path = rtweet_stream_path2)
```


Fortunately, `{tweetio}` can handle these situations by validating the JSON before it gets anywhere near R.

```{r}
bench::mark(
  read_tweets(rtweet_stream_path),
  min_iterations = 3
)[, 1:9]
```


Ignoring the time it takes to run the above `rtweet_validate_and_parse()` function for `rtweet::parse_stream()` to work, how much faster is `tweetio::read_tweets()`?

```{r, out.width="100%"}
bench_mark <- bench::mark(
  rtweet = rtweet::parse_stream(rtweet_stream_path2),
  tweetio = tweetio::read_tweets(rtweet_stream_path2),
  check = FALSE,
  min_iterations = 3,
  filter_gc = FALSE
)

bench_mark[, 1:9]

plot(bench_mark)
```


With bigger files, using `rtweet::parse_stream()` is no longer realistic, but that's where `tweetio::read_tweets()` can help.

```{r}
big_tweet_stream_path <- "~/ufc-tweet-stream.json.gz"
scales::number_bytes(file.size(big_tweet_stream_path)) # 1.2 Gb before compression

bench::mark(
  big_tweet_df <- tweetio::read_tweets(big_tweet_stream_path)
)[, 1:9]
```



<!-- ## Data Dumps -->

<!-- Using Elasticsearch seems to be the most common practice for handling social media data at scale, but it's (unfortunately) possible that you'll need to work with data dumps. -->

<!-- The data collected from APIs are stored inside a nested JSON object named `"doc"`. `"doc"` is typically embedded in a ton of system metadata, most of which you probably don't care about.  -->

<!-- I've encountered two flavors of these schema: -->

<!-- 1. .jsonl: newline-delimited JSON -->
<!-- 2. .json: the complete contents of a database dump packed in a JSON array -->

<!-- The .jsonl structure looks something like this: -->

<!-- ```{r, eval=FALSE, echo=FALSE} -->
<!-- es_jsonl <- readLines("inst/example-data/elasticsearch-dump-example.jsonl") -->

<!-- cat(paste0(es_jsonl, collapse = "\n")) -->
<!-- ``` -->

<!-- ```json -->
<!-- {"system_metadata_you_dont_care_about":"blahblahblah","more_metadata1":"blahblahblah","more_metadata2":"blahblahblah","more_metadata3":"blahblahblah","more_metadata4":"blahblahblah","more_metadata5":"blahblahblah","doc":{"text":"********************HERE'S THE DATA YOU ACTUALLY WANT********************","id":92108498098018010401,"id_str":"92108498098018010401"},"more_metadata6":"blahblahblah","more_metadata7":"blahblahblah","more_metadata8":"blahblahblah","more_metadata9":"blahblahblah","more_metadata10":"blahblahblah"} -->
<!-- {"system_metadata_you_dont_care_about":"blahblahblah","more_metadata1":"blahblahblah","more_metadata2":"blahblahblah","more_metadata3":"blahblahblah","more_metadata4":"blahblahblah","more_metadata5":"blahblahblah","doc":{"text":"********************HERE'S THE DATA YOU ACTUALLY WANT********************","id":92108498098018010401,"id_str":"92108498098018010401"},"more_metadata6":"blahblahblah","more_metadata7":"blahblahblah","more_metadata8":"blahblahblah","more_metadata9":"blahblahblah","more_metadata10":"blahblahblah"} -->
<!-- {"system_metadata_you_dont_care_about":"blahblahblah","more_metadata1":"blahblahblah","more_metadata2":"blahblahblah","more_metadata3":"blahblahblah","more_metadata4":"blahblahblah","more_metadata5":"blahblahblah","doc":{"text":"********************HERE'S THE DATA YOU ACTUALLY WANT********************","id":92108498098018010401,"id_str":"92108498098018010401"},"more_metadata6":"blahblahblah","more_metadata7":"blahblahblah","more_metadata8":"blahblahblah","more_metadata9":"blahblahblah","more_metadata10":"blahblahblah"} -->
<!-- ``` -->

<!-- Each line contains a single JSON object resembling the following: -->

<!-- ```{r, echo=FALSE, eval=FALSE} -->
<!-- jsonlite::prettify(es_jsonl[[1]]) -->
<!-- ``` -->

<!-- ```json -->
<!-- { -->
<!--     "system_metadata_you_dont_care_about": "blahblahblah", -->
<!--     "more_metadata1": "blahblahblah", -->
<!--     "more_metadata2": "blahblahblah", -->
<!--     "more_metadata3": "blahblahblah", -->
<!--     "more_metadata4": "blahblahblah", -->
<!--     "more_metadata5": "blahblahblah", -->
<!--     "doc": { -->
<!--         "text": "********************HERE'S THE DATA YOU ACTUALLY WANT********************", -->
<!--         "id": 92108498098018010401, -->
<!--         "id_str": "92108498098018010401" -->
<!--     }, -->
<!--     "more_metadata6": "blahblahblah", -->
<!--     "more_metadata7": "blahblahblah", -->
<!--     "more_metadata8": "blahblahblah", -->
<!--     "more_metadata9": "blahblahblah", -->
<!--     "more_metadata10": "blahblahblah" -->
<!-- } -->
<!-- ``` -->

<!-- And the .json structure looks something like this: -->


<!-- ```{r, echo=FALSE, eval=FALSE} -->
<!-- es_json <- readr::read_file("inst/example-data/elasticsearch-dump-example.json") -->
<!-- jsonlite::prettify(es_json) -->
<!-- ``` -->

<!-- ```json -->
<!-- [ -->
<!--     { -->
<!--         "_id": "e5daf1467d2438e31b11b44a82cbd7f5758ba5a1f1d3ecbcc6e1fc04dc9c7c4d-3016858092318", -->
<!--         "_index": "org-77f135f331153568ab7eb0e4c24623a7-default-3769a33b9e88598e38317591e2ee31c3-default-030009", -->
<!--         "_score": null, -->
<!--         "_source": { -->
<!--             "system_metadata_you_dont_care_about": "blahblahblah", -->
<!--             "more_metadata1": "blahblahblah", -->
<!--             "doc": { -->
<!--                 "text": "********************HERE'S THE DATA YOU ACTUALLY WANT********************", -->
<!--                 "id": 92108498098018010401, -->
<!--                 "id_str": "92108498098018010401" -->
<!--             }, -->
<!--             "more_metadata6": "blahblahblah" -->
<!--         } -->
<!--     }, -->
<!--     { -->
<!--         "_id": "e5daf1467d2438e31b11b44a82cbd7f5758ba5a1f1d3ecbcc6e1fc04dc9c7c4d-3016858092318", -->
<!--         "_index": "org-77f135f331153568ab7eb0e4c24623a7-default-3769a33b9e88598e38317591e2ee31c3-default-030009", -->
<!--         "_score": null, -->
<!--         "_source": { -->
<!--             "system_metadata_you_dont_care_about": "blahblahblah", -->
<!--             "more_metadata1": "blahblahblah", -->
<!--             "doc": { -->
<!--                 "text": "********************HERE'S THE DATA YOU ACTUALLY WANT********************", -->
<!--                 "id": 92108498098018010401, -->
<!--                 "id_str": "92108498098018010401" -->
<!--             }, -->
<!--             "more_metadata6": "blahblahblah" -->
<!--         } -->
<!--     } -->
<!-- ] -->
<!-- ``` -->

<!-- This has three unfortunate consequences: -->

<!-- 1. Packages that were purpose-built to work directly with `{rtweet}`'s data frames can't play along with your data. -->
<!-- 2. You're going to waste most of your time (and memory) getting data into R that you're not going to use. -->
<!-- 3. The data are _very_ tedious to restructure in R (lists of lists of lists of lists of lists...). -->

<!-- `{tweetio}` solves this by parsing everything and building the data frames at the C++ level. -->



# Spatial Tweets


Tweet JSON contains an object called `"place"` that is set aside for location information. If present, it looks something like the following.

```json
{
    "text": "#rstats4lyfe",
    "place": {
      "id": "00d546b224a6764d",
      "url": "https://api.twitter.com/1.1/geo/id/blahblah.json",
      "place_type": "city",
      "name": "General Pico",
      "full_name": "General Pico, Argentina",
      "country_code": "AR",
      "country": "Argentina",
      "bounding_box": {
          "type": "Polygon",
          "coordinates": [
              [
                  [
                      -63.784544,
                      -35.689402
                  ],
                  [
                      -63.784544,
                      -35.641737
                  ],
                  [
                      -63.724717,
                      -35.641737
                  ],
                  [
                      -63.724717,
                      -35.689402
                  ]
              ]
          ]
      },
      "attributes": {}
    }
}
```

In `{rtweet}`-style data frames, this corresponds to the `bbox_coords` column.

If you have `{sf}` installed, you can use `as_tweet_sf()` to filter the tweets for those that contain valid coordinates and build their bounding box polygons.


```{r}
tweet_sf <- as_tweet_sf(tweet_df = big_tweet_df) 

tweet_sf[, "bbox_coords"]
```


There are currently three columns that can potentially hold bounding boxes:

1. `bbox_coords`
2. `quoted_bbox_coords`
3. `retweet_bbox_coords`

You can select which one to use to build your `sf` object by modifying the `geom_col=` parameter (default: `"bbox_coords"`)

```{r}
as_tweet_sf(big_tweet_df, geom_col = "quoted_bbox_coords"
            )[, "quoted_bbox_coords"]
```

You can also build _all_ the supported bounding boxes by setting `geom_col=` to `"all"`.

```{r}
all_bboxes <- as_tweet_sf(big_tweet_df, geom_col = "all")
all_bboxes[, c("geometry", "which_geom")]
unique(all_bboxes$which_geom)
```


From there, you can easily use the data like any other `{sf}` object.


```{r, fig.width=12, fig.height=8}
library(ggplot2)

world <- rnaturalearth::ne_countries(returnclass = "sf")
world <- world[world$continent != "Antarctica", ]

sf_for_gg <- sf::st_wrap_dateline(all_bboxes)

ggplot(sf_for_gg) +
  geom_sf(fill = "white", color = "lightgray", data = world) +
  geom_sf(aes(fill = which_geom, color = which_geom), alpha = 0.15, size = 1, show.legend = TRUE) +
  coord_sf(crs = 3857) +
  scale_fill_viridis_d() +
  scale_color_viridis_d() +
  theme(legend.title = element_blank(), legend.position = "top",
        panel.background = element_rect(fill = "#daf3ff"))
```



# Can we go faster?

Until Rtools 4.0 hits (or the [`simdjson`](https://github.com/lemire/simdjson) library decides to [relax its C++17 requirement](https://github.com/lemire/simdjson/issues/307)), I'm not sure how we can go much faster while maintaining cross-platform compatibility. That said, contibutions are welcome, especially if C++ is your mother tongue.



# Acknowledgements

`{tweetio}` uses a combination of C++ via [`{Rcpp}`](http://www.rcpp.org/), the [`rapidjson`](http://rapidjson.org/) C++ library (made available by [`{rapidjsonr}`](https://cran.r-project.org/web/packages/rapidjsonr/index.html)), [`{jsonify}`](https://cran.r-project.org/web/packages/jsonify/index.html)) for an R-level interface to `rapidjson`, [`{RcppProgress}`](https://cran.r-project.org/web/packages/RcppProgress/index.html)), and __R's secret weapon__: [`{data.table}`](https://rdatatable.gitlab.io/data.table/).

Major inspiration from [`{ndjson}`](https://gitlab.com/hrbrmstr/ndjson) was taken, particularly its use of [`Gzstream`](https://www.cs.unc.edu/Research/compgeom/gzstream/).
